---
title: "Homework 2"
author: Alisha Sarakki
output: github_document
---
```{r}
library(tidyverse)
library(readxl)
library(dplyr)
tibble()
```

## Problem 1

Read the Mr. Trash-Wheel dataset.

```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  ) 

```


Read precipitation data!

```{r}
precip_2018 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
```
```{r}
precip_2017 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)

```

Now combine annual precipitation.

```{r}

month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018, precip_2017)
  
left_join(precip_df, month_df, by = "month")

```


This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, MD. As the trash enters the inner harbor, the trashwheel collects that trash adn stores it in a dumpster. The dataset contains information on year, month, trash collected, and specific types of trash. There are a total of 'r nrow(trashwheel_df)' rows in our final dataset. Additional data sheets include month precipitation data.


## Problem 2

Import and clean NYC transit data

```{r}
NYCTransit_data = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
  names(NYCTransit_data)
  NYCTransit_data = janitor::clean_names(NYCTransit_data)
  names(NYCTransit_data)

```

Specify which columns to keep

```{r}
NYCT1_data = 
  select(NYCTransit_data, line:entry, vending, ada)
```

Character to Logical Expression Conversion
```{r}
NYCT1_data %>% 
  mutate(entry = recode(entry, YES = "TRUE", NO = "FALSE"))
```

This dataset describes information about each entrance and exit for all subway stations in New York, including, broadly, detailed location of both station and entrance/exit, station name, which routes they lie on, descriptions of entrances (stairs/escalator/etc.), and ADA compliance. Thus far, I have narrowed the scope of this data to include line, latitude and longitude, station name, train routes, entrance type, and whether the station is ADA compliant. I have also cleaned the names of the dataset, and converted the entry expression of "YES/NO" to a logical expression of "TRUE/FALSE".The dimensions are `r nrow(NYCT1_data)` x `r ncol(NYCT1_data)`.These data are not tidy.



Distinct Stations

```{r}
  distinct(NYCT1_data, station_name, line)
count(distinct(NYCT1_data, station_name, line))
```
465 total distinct stations

ADA Compliant Stations
```{r}
filter(NYCT1_data, ada == TRUE) %>% 
distinct(line, station_name) %>%
count()
```
84 stations are ADA compliant. 

Proportion of station entrances/exits without vending that allow entrance
```{r}
NYCT1_data %>% 
  mutate(entry = recode(entry, YES = "TRUE", NO = "FALSE")) %>% 
filter(vending == "NO", entry == "TRUE") %>% 
  count()

NYCT1_data %>% 
  mutate(entry = recode(entry, YES = "TRUE", NO = "FALSE")) %>% 
filter(vending =="NO") %>% 
count()

69/183
```
69 total station entries and exits without vending that allow entrance
183 total station entries and exits without vending
Final proportion: 0.3770492

Reformat route number and route name
```{r}
NYCT_tidy =
  NYCT1_data%>%
  mutate_at(vars(route1:route11), as.character) %>%
  pivot_longer(
    route1:route11,
    names_to = "route_no",
    values_to = "route_name",
    names_prefix = "route",
  )
```

Distinct Stations Serving A train
```{r}
filter(NYCT_tidy, route_name == "A" ) %>% 
distinct(line, station_name)

```
There are `r nrow(NYCT_tidy)` distinct stations that serve the A line. 


```{r}
filter(NYCT_tidy, route_name == "A", ada == TRUE) %>% 
distinct(line, station_name) %>%
count()
```
Of all distinct stations that serve the A, 17 are ADA compliant.


## Problem 3

Import 538 data
```{r}
pols_data = 
  read_csv(file = "./data/pols-month.csv")
snp_data = 
  read_csv(file = "./data/snp.csv")
unemployment_data = 
    read_csv(file = "./data/unemployment.csv" )
```

Clean pols-month data
```{r}
clean_pols = 
  pols_data %>% 
  separate("mon", into = c("year", "month", "day")) %>% 
  mutate(month = as.numeric(month))

month_pols = 
  tibble( 
   month = (1:12),
    month_name = month.name) %>% 
    mutate(months = as.numeric(month))

clean_month_pols = left_join(clean_pols, month_pols, by = "month")

clean_month_pols %>% 
  mutate(president = case_when(
    prez_gop == 1 ~ "gop",
    prez_dem == 1 ~ "dem")) %>% 
    select(-prez_gop, -prez_dem, -month, -day)

head(arrange(clean_month_pols, year, month_name))

intermed_pols_ = 
  clean_month_pols %>% 
  mutate(year = as.numeric(year))

```


snp cleaning + tidying

```{r}
clean_snp = 
  snp_data %>% 
  separate("date", into = c("month", "day", "year")) %>% 
  mutate(month = as.numeric(month))

month_snp = 
  tibble( 
   month = (1:12),
    month_name = month.name) %>% 
    mutate(month = as.numeric(month))

clean_month_snp = left_join(clean_snp, month_snp, by = "month")

clean_month_snp %>% 
  relocate(year, month)

intermed_snp = 
  clean_month_snp %>% 
  mutate(year = as.numeric(year))


```

Tidy unemployment data
```{r}
clean_unemp = 
  unemployment_data %>% 
  janitor::clean_names() %>% 
 pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "percent unemployed"
 ) %>% 
   mutate(month = as.numeric(month))
  
intermed_unemploy = 
  clean_unemp %>% 
  mutate((month = recode(month, jan = "january", feb = "february", mar = "march", apr = "april", may = "may", jun = "june", jul = "july", aug = "august", sep = "september", oct = "october", nov = "november", dec = "december")))
 
intermed_unemploy_2 = 
  clean_unemp %>% 
  mutate(year = as.numeric(year))

month_unemp = 
  tibble( 
   month = (1:12),
    month_name = month.name) %>% 
    mutate("month" = as.numeric(month))

final_unemploy =
  left_join(intermed_unemploy, month_unemp)

```


Joining datasets
```{r}
pols_snp = 
  left_join(intermed_pols_, intermed_snp)

final_join =  
  left_join(pols_snp, intermed_unemploy)

final_join
```
Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables

These data describe the number of national politicians that are democratic or republican at any given time, data pertaining to the stock market (S&P index), and percentage/proportion of unemployment at a given time. The dimensions are `r nrow(final_join)` x `r ncol(final_join)`. The range of years is between `r min(pull(final_join))` and `r max(pull(final_join))`.  Key variables include governor/senator and presidential political affiliation, as related to time.  


