---
title: "Homework 2"
author: Alisha Sarakki
output: github_document
---
```{r}
library(tidyverse)
library(readxl)
library(dplyr)
tibble()
```

## Problem 1

Read the Mr. Trash-Wheel dataset.

```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  ) 

```


Read precipitation data!

```{r}
precip_2018 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
```
```{r}
precip_2017 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)

```

Now combine annual precipitation.

```{r}

month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018, precip_2017)
  
left_join(precip_df, month_df, by = "month")

```


This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, MD. As the trash enters the inner harbor, the trashwheel collects that trash adn stores it in a dumpster. The dataset contains information on year, month, trash collected, and specific types of trash. There are a total of 'r nrow(trashwheel_df)' rows in our final dataset. Additional data sheets include month precipitation data.


## Problem 2

Import and clean NYC transit data

```{r}
NYCTransit_data = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
  names(NYCTransit_data)
  NYCTransit_data = janitor::clean_names(NYCTransit_data)
  names(NYCTransit_data)

```

Specify which columns to keep

```{r}
select(NYCTransit_data, line:entry, vending, ada)
```

Character to Logical Expression Conversion
```{r}
NYCTransit_data %>% 
  mutate(entry = recode(entry, YES = "TRUE", NO = "FALSE"))
```

This dataset describes information about each entrance and exit for all subway stations in New York, including, broadly, detailed location of both station and entrance/exit, station name, which routes they lie on, descriptions of entrances (stairs/escalator/etc.), and ADA compliance. Thus far, I have narrowed the scope of this data to include line, latitude and longitude, station name, train routes, entrance type, and whether the station is ADA compliant. I have also cleaned the names of the dataset, and converted the entry expression of "YES/NO" to a logical expression of "TRUE/FALSE".The dimensions are 'nrow(NYCTransit_data)' x 'ncol(NYCTransit_data)'.These data are not tidy.

```{r}
nrow(NYCTransit_data)
ncol(NYCTransit_data)
```

Distinct Stations

```{r}
  distinct(NYCTransit_data, station_name, line)
count(distinct(NYCTransit_data, station_name, line))
```
465 total distinct stations

ADA Compliant Stations
```{r}
filter(NYCTransit_data, ada == TRUE) %>% 
distinct(line, station_name) %>%
count()
```
84 stations are ADA compliant. 

Proportion of station entrances/exits without vending that allow entrance
```{r}
NYCTransit_data %>% 
  mutate(entry = recode(entry, YES = "TRUE", NO = "FALSE")) %>% 
filter(vending == "NO", entry == "TRUE") %>% 
  count()

NYCTransit_data %>% 
  mutate(entry = recode(entry, YES = "TRUE", NO = "FALSE")) %>% 
filter(vending =="NO") %>% 
count()

69/183
```
69 total station entries and exits without vending that allow entrance
183 total station entries and exits without vending
Final proportion: 0.3770492

Reformat route number and route name
```{r}
NYCT_tidy =
  NYCTransit_data%>%
  mutate_at(vars(route1:route11), as.character) %>%
  pivot_longer(
    route1:route11,
    names_to = "route_no",
    values_to = "route_name",
    names_prefix = "route",
  )
```

Distinct Stations Serving A train
```{r}
filter(NYCT_tidy, "route_name" == "A" ) %>% 
distinct(line, station_name) %>%
count()
```

Of all distinct stations that serve the A, the following number are ADA compliant.
```{r}
filter(NYCT_tidy, "route_name" == "A", ada == TRUE) %>% 
distinct(line, station_name) %>%
count()
```

## Problem 3

Import 538 data
```{r}
pols_data = 
  read_csv(file = "./data/pols-month.csv")
snp_data = 
  read_csv(file = "./data/snp.csv")
unemployment_data = 
    read_csv(file = "./data/unemployment.csv" )
```

Clean pols-month data
```{r}
clean_pols = 
  pols_data %>% 
  separate("mon", into = c("year", "month", "day")) %>% 
  mutate(month = as.numeric(month))

month_pols = 
  tibble( 
   month = (1:12),
    month_name = month.name) %>% 
    mutate(month = as.numeric(month))

clean_month_pols = left_join(clean_pols, month_pols, by = "month")

clean_month_pols %>% 
  mutate(president = case_when(
    prez_gop == 1 ~ "gop",
    prez_dem == 1 ~ "dem")) %>% 
    select(-prez_gop, -prez_dem, -month, -day)

head(arrange(clean_month_pols, year, month_name))
  
```

snp cleaning + tidying

```{r}
snp_data
clean_snp = 
  snp_data %>% 
  separate("date", into = c("month", "day", "year")) %>% 
  mutate(month = as.numeric(month))

month_snp = 
  tibble( 
   month = (1:12),
    month_name = month.name) %>% 
    mutate(month = as.numeric(month))

clean_month_snp = left_join(clean_snp, month_snp, by = "month")

clean_month_snp %>% 
  relocate(year, month)
```

